A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-10-02 kobakian_cook.Rmd
Potentially use NSW for thyroid example?

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, comment = FALSE}
library(knitcitations)
library(RefManageR)
library(sf)
library(sugarbag)
library(kableExtra)

knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options("citation_format" = "pandoc")
BibOptions(check.entries = FALSE, style = "markdown")
```

@@

<!-- Issues with choropleth map displays... -->
In many countries, residents are increasingly gathering to live in urban areas around major cities [@ACTUC].
The creation of approximately equal population areas results in dramatically different square meterage of the geographic units. The geographic units are usually administrative, such as states or electorates. The units are filled with colour to represent the value of the statistic for each unit [@EI].
This can lead to unequal emphasis on the statistical value of each of the geographic units. This allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM].
Figure \ref{fig:melanoma-geo} is a choropleth map that uses colour to display the estimated Standardised Incidence Ratios (SIRs) of melanoma cancer for males for each of the Statistical Areas at Level 2 (SA2) used by the Australian Bureau of Statistics [-@abs2011]. The Australian choropleth map display draws attention to the expanse of light blue areas across the rural communities in all states. The SA2s around Brisbane stand out as more orange and red. To create this choropleth map the SA2 polygons for 2011 from the Australian Bureau of Statistics.

@@

There are several established alternative visualisation methods. 
Tile maps, Rectangular cartograms [@ORC] and Dorling cartograms [@ACTUC] all use one simple shape to represent each geographic unit. They all minimise the emphasis on the size or shape of the geographic areas. These alternative map displays focus on the relationship between neighbours, attempting to preserve connections, and disregard the unique shapes of the administrative boundaries. Figure \ref{fig:qld_grid} shows a collection of alternative map displays, this includes a) a contiguous cartogram, b) a non-contiguous cartogram, c) a Dorling cartogram and d) a hexagon tile map. 
Each Statistical Area at Level 2 (SA2) [@abs2011] was designed to represent a community. These administrative areas are often used to aggregate census data and used to understand demographics of the Australian population.

<!-- choropleth -->
The familiar choropleth map display is often used to allow the user to orient themselves. However, the emphasis on the land mass diminishes the features of the distribution in densely populated communities due to the small size on the display [@ACTUC].
The unique shapes of boundaries can be helpful for orienting users but may not contribute to their understanding of the spatial disease distribution as many of the communities are not visible in a choropleth display [@TVSSS]. This presents an opportunity for explicit map transformations to improve communication of distributions [@CBATCC]. In figure \ref{fig:qld_grid}a) the choropleth map is presented with the alternative maps to allow for comparisons to be made.

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/qld_grid.pdf}
\caption{\label{fig:qld_grid}A collection of alternative maps of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Melanoma for males, ranging from much lower than average (blue) to much higher than average (red)}
\end{figure}

<!-- contiguous -->
When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS].
The impact of a disease over connected communities can be seen by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in figure \ref{fig:qld_grid}b) draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS]. This shown in the northern SA2s of Queensland with an average (yellow) estimated SIR of Melanoma for males.

<!-- non-contiguous -->
The non-contiguous cartogram in figure \ref{fig:qld_grid}c) also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of white space can be meaningful in non-contiguous cartograms [@ECGC], in this example the disparity between large rural areas and small urban areas means that the city of Brisbane, shown in the inset map looks reasonable, but these units are not visible in the context of the Queensland state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

<!-- Dorling -->
The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \ref{fig:qld_grid}d) shows the Queensland SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in Queensland that were not apparent in figure \ref{fig:qld_grid} a), b) or c).

@@

The following code creates the hexagon tile map for all the Statistical Areas at Level 2 in Queensland. These steps can be executed by the main function, `create_hexmap`, or can be implemented separately for more flexibility.

@@

```{r full_hexmapcode, echo = TRUE, warning = FALSE, message = FALSE, eval=FALSE}
## Load data
qld_sa2 <- absmapsdata::sa22011 %>% 
  filter(state_name_2011 == "Queensland")

## Create centroids set
centroids <- create_centroids(qld_sa2, "sa2_name_2011")

## Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.12,
  ## same size used in create_grid
  hex_filter = 10,
  use_neighbours = qld_sa2,
  focal_points = capital_cities %>%
    filter(points == "Brisbane"),
  width = 35,
  verbose = FALSE
)

fort_qld <- qld_sa2 %>%
  fortify_sfc()

## Make a plot
library(ggplot2)
qld_hexmap <- ggplot() +
  geom_polygon(aes(
      x = long,
      y = lat,
      group = interaction(sa2_name_2011, polygon)
    ),
    fill = "white",
    colour = "lightgrey",
    data = fort_qld
  ) +
  geom_polygon(aes(
    x = long,
    y = lat,
    group = hex_id,
    fill = rownumber
  ),
  data = fort_hex) +
  scale_fill_distiller("", palette = "PRGn") +
  coord_equal() + theme_void()
qld_hexmap
```

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/aus_gghexmap.pdf}
\caption{\label{fig:melanoma-hex}A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR, they range from much lower than average (blue) to much higher than average (red)}
\end{figure}

@@

The polygon set of Statistical Areas at Level 2 (SA2) [@abs2011] of Queensland in 2011 is provided with the `sugarbag` package as `tas_sa2`. 
A single column of the data set is used to identify the unique areas. 
In this case, the unique SA2 names for each SA2 have been used.

The longitude and latitude centre of the capital cities of Australia are used as focal points to allocate each geographic area around the closest capital city. Brisbane will be the common focal point, as this example uses only the areas in the state of Queensland.

@@

<!-- ```{r grid2, fig.cap = "Grid points to create a tile map.", out.width = "100%", echo = FALSE} -->
<!-- knitr::include_graphics(path = "figures/2grid.pdf") -->
<!-- ``` -->

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/2grid.pdf}
\caption{\label{fig:grid2}Grid points to create a hexagon tile map.}
\end{figure}

@@

The grid points are kept only if they fall between the rolling average of the minimum and maximum centroid values after accounting for the buffer distance, for each row and column of the grid. The sparsely populated South-West region of National Park has much fewer points available compared to the South-East region containing the city of Brisbane.

<!-- ```{r filter-grid, fig.cap = "All possible hexagon locations from the initial grid are shown with blue outlines. The blue dots show the grid points left to choose from after the buffer step. The rolling windows show the collections of rows and columns used to filter the hexagon locations.", echo = FALSE, out.width = "100%"} -->
<!-- knitr::include_graphics(path = "figures/3grid.pdf", dpi = 300) -->
<!-- ``` -->

\begin{figure}[h]
\centering
\includegraphics[width=10cm]{figs/3grid.pdf}
\caption{\label{fig:filter-grid}The remaining hexagon locations from all the possible points in the initial grid after the buffer is applied. The blue dots show the grid points left to choose from after the buffer step.}
\end{figure}

@@

<!-- ```{r buffers, fig.cap = "Filter for grid points within a square, then circular, distance for those closest to the centroid.", echo = FALSE, out.width = "100%"} -->
<!-- knitr::include_graphics(path = "figures/4grid.pdf") -->
<!-- ``` -->

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/4grid.pdf}
\caption{\label{fig:buffers}The remaining available hexagon locations after filtering for the grid points within a square distance, then circular around the centroid.}
\end{figure}

@@

<!-- ```{r buffs, fig.cap = "A complete hexagon tile map of Queensland.", out.width = "100%", echo = FALSE} -->
<!-- knitr::include_graphics(path = "figures/6allocate.png") -->
<!-- ``` -->

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/6allocate.pdf}
\caption{\label{fig:buffs}A complete hexagon tile map of Queensland.}
\end{figure}

A complete hexagon tile map of Queensland is created by applying the algorithm steps to each centroid.
The hexagon tile map visualisation is used below to visualise the Australian Cancer Atlas data [@TACA]. Two views of the same data are produced by filling according to the Lung Cancer Standardised Incidence Rates (SIRs) downloaded from the Australian Cancer Atlas site. This small example in Figure \ref{fig:sir} shows the group of blue areas in the Brisbane CBD more prominently in the hexagon tile map (b).
The small red areas visible in the choropleth map (a) along the north coast are much larger in the hexagon tile maps. The hexagon tile map shows less yellow, this no longer overwhelms the map space with the information regarding the rural areas.

@@

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/7SIR.pdf}
\caption{\label{fig:sir}The Australian Cancer Atlas data displayed on a choropleth and hexagon tile map. The colour of each Statistical Area of Australian at Level 2 coloured by SIR. A choropleth map (a) of SIR is paired with a hexagon tile map (b) to contrast the colours that are made obvious when every SA2 is equally represented.}
\end{figure}

@@

The `gganimate` [@gganimate] package can be used to make an animation.
It requires connecting the polygons for each area in two displays, which can be done using the `sf_id` variable, such as the SA2 name.
The animation^[This animation can be viewed at: https://sugarbagjss.netlify.com/] connecting these two displays will highlight the rapid growth of the inner-city areas that emphasises the density of the communities. The hexagons that move the furthest will move rapidly in the animation. The rapid decreases of the large rural areas also show how much greater the landmass of Statistical Areas can be.

@@

Alternative maps highlight the value of the statistics across the geographic units. Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing each administrative area. This acknowledges that the amount of residents can be different but recognises that each population area is equally important. The solution to this visualisation problem has equally sized areas, with neighbourhood boundary connections. This map algorithm is implemented in the `sugarbag` [@sugarbag] package written for `R` [@R].
The `sugarbag` package creates tessellated hexagon tile maps. The Australian application preserves the spatial relationships, emphasising capital cities. The hexagon tile map is a visualisation solution that highlights spatial distributions.

These hexagons equally represent each area. However, the tessellation does not allow the size of the hexagons to represent another variable, similar to the choropleth maps.
The algorithm is heavily dependent on the focal points used, as this determines the order of allocation. It works on the assumption that viewers can use directional relationships to identify their neighbourhoods but this can be aided by the animation.
With careful consideration of the choropleth map, the small geographic inner city areas may have been noticed by viewers, but the hexagon tile map display emphasises them. The communities in northern Queensland and the Northern territory do not draw attention because of their size as in the choropleth, but their colour is still noticeably below average when contrasted with the hexagons further south.

<!-- why hexagon tile map addresses deficiencies -->
The Dorling cartogram was the most influential in the design of the hexagon tile map.
The use of a hexagon provides a simple shape that shifts the focus of the visualisation from the shape of SA2s to the distribution effects of a disease on communities.  not impeded by the irregular and unusual shapes created by a contiguous cartogram and the immense white space created on the non-contiguous cartogram.
Hexagons can emphasise the relationship between neighbouring communities by allowing their sides to touch. The tessellation crated by adjoining hexagons for densely populated areas is helpful as it shows that most communities are not islands onto themselves, but are intrinsically connected to those around them. As seen in the Dorling cartogram in figure \ref{fig:qld_grid}c), the tessellation that allows boundary connections was not imposed for rural communities, where centroid of the geographic unit was located far from others, a feature that allows distant but dense populations of the coastal towns of Queensland to be recognised.

>>>>

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, comment = FALSE}
library(knitcitations)
library(kableExtra)
library(RefManageR)
library(sp)
library(sf)
library(sugarbag)
library(absmapsdata)
library(tidyverse)
library(cartogram)
library(ggthemes)
library(cowplot)

knitr::opts_chunk$set(warning = FALSE, message = FALSE, eval = FALSE)
options("citation_format" = "pandoc")
BibOptions(check.entries = FALSE, style = "markdown")

invthm <- theme_map() + 
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA), 
    plot.background = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.key = element_rect(fill = "transparent", colour = NA),
    text = element_text(colour = "white"),
    axis.text = element_text(colour = "white")
  )

# function to allocate colours to regions
aus_colours <- function(sir_p50){
  value <- case_when(
    sir_p50 <  0.74 ~ "#33809d",
    sir_p50 >= 0.74 & sir_p50 < 0.98 ~ "#aec6c7",
    sir_p50 >= 0.98 & sir_p50 < 1.05 ~ "#fff4bc",
    sir_p50 >= 1.05 & sir_p50 < 1.45 ~ "#ff9a64",
    sir_p50 >= 1.45 ~ "#ff3500",
    TRUE ~ "#FFFFFF")
  return(value)
}

```

@@

<!-- Issues with choropleth map displays... -->
In many countries, residents are increasingly gathering to live in urban areas around major cities [@ACTUC].
The creation of approximately equal population areas results in dramatically different square meterage of the administrative geographic units, such as states or electorates. The units are filled with colour to represent the value of the statistic for each unit [@EI].
This can lead to unequal emphasis on the statistical value of each of the geographic units. This allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM].
Figure \ref{fig:melanoma-geo} is a choropleth map that uses colour to display the estimated Standardised Incidence Ratios (SIRs) of melanoma cancer for males for each of the Statistical Areas at Level 2 (SA2) used by the Australian Bureau of Statistics [-@abs2011]. The Australian choropleth map display draws attention to the expanse of light blue areas across the rural communities in all states. The SA2s around Hobart stand out as more orange and red. To create this choropleth map the SA2 polygons for 2011 from the Australian Bureau of Statistics.

@@

There are several established alternative visualisation methods. 
Tile maps, Rectangular cartograms [@ORC] and Dorling cartograms [@ACTUC] all use one simple shape to represent each geographic unit. They all minimise the emphasis on the size or shape of the geographic areas. These alternative map displays focus on the relationship between neighbours, attempting to preserve connections, and disregard the unique shapes of the administrative boundaries. Figure \ref{fig:nsw_grid} shows a collection of alternative map displays, this includes a) a contiguous cartogram, b) a non-contiguous cartogram, c) a Dorling cartogram and d) a hexagon tile map. 
Each Statistical Area at Level 2 (SA2) [@abs2011] was designed to represent a community. These administrative areas are often used to aggregate census data and used to understand demographics of the Australian population.

<!-- choropleth -->
The familiar choropleth map display is often used to allow the user to orient themselves. However, the emphasis on the land mass diminishes the features of the distribution in densely populated communities due to the small size on the display [@ACTUC].
The unique shapes of boundaries can be helpful for orienting users but may not contribute to their understanding of the spatial disease distribution as many of the communities are not visible in a choropleth display [@TVSSS]. This presents an opportunity for explicit map transformations to improve communication of distributions [@CBATCC]. In figure \ref{fig:nsw_grid}a) the choropleth map is presented with the alternative maps to allow for comparisons to be made.

```{r}
states <- absmapsdata::state2011 %>% 
  rmapshaper::ms_simplify(keep = 0.25, keep_shapes = TRUE) %>% 
  filter(!state_name_2011 == "Other Territories")

sa2 <- absmapsdata::sa22011 %>% 
  filter(state_name_2011 == "New South Wales") %>% 
  filter(!st_is_empty(geometry))

# sa2 <- sa2 %>% rmapshaper::ms_simplify(keep = 0.05, keep_shapes = TRUE)

SIR <- read_csv("data/SIR Downloadable Data.csv") %>% 
  filter(SA2_name %in% sa2$sa2_name_2011) %>% 
  dplyr::select(Cancer_name, SA2_name, Sex_name, p50) %>% 
  filter(Cancer_name == "Thyroid", Sex_name == "Females")
ERP <- read_csv("data/ERP.csv") %>%
  filter(REGIONTYPE == "SA2", Time == 2011, Region %in% SIR$SA2_name) %>% 
  dplyr::select(Region, Value)

# Alternative maps
# Join with sa2 sf object
sa2thyroid_ERP <- SIR %>% 
  left_join(sa2, ., by = c("sa2_name_2011" = "SA2_name")) %>%
  left_join(., ERP %>% 
              dplyr::select(Region, 
              Population = Value), by = c("sa2_name_2011"= "Region")) %>% 
  filter(!st_is_empty(geometry))

sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf()

aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  #geom_sf(fill = "lightgrey", data = states) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity()
```

```{r alternatives}

###############################################################################

sa2thyroid_ERP <- st_transform(sa2thyroid_ERP, 3112)
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  mutate(sva = sqrt(as.numeric(Population)/as.numeric(albers_sqkm)))

###############################################################################
# Contiguous Cartograms

cont <- sa2thyroid_ERP %>% 
  mutate(Population = Population + 1) %>% 
  cartogram_cont(., weight = "Population", itermax = 20) %>%
  st_as_sf()

save(cont, file = "data/cont.rda")
load("data/cont.rda")

aus_ggcont <- ggplot(cont) + 
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
aus_ggcont

ggsave(filename = "figures/aus_ggcont.png", plot = aus_ggcont,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

```

```{r}

###############################################################################
# Non - Contiguous Cartograms

ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1/10,
                         weight = "Population") %>% st_as_sf()

save(ncont, file = "data/ncont.rda")
load("data/ncont.rda")

aus_ggncont <- ggplot(ncont) + 
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
aus_ggncont

ggsave(filename = "figures/aus_ggncont.png", plot = aus_ggncont,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

```

```{r}

###############################################################################
# Non - Contiguous Dorling Cartograms

dorl <- sa2lung_ERP %>%
  mutate(pop = (Population/max(Population))*10) %>% 
  cartogram_dorling(., k = 0.01, weight = "pop", m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
aus_ggdorl <- ggplot(dorl) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity()+
  invthm + guides(fill = FALSE)
aus_ggdorl

ggsave(filename = "figures/aus_ggdorl.png", plot = aus_ggdorl,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

```

```{r full_hexmapcode}
## Load data
tas_sa2 <- absmapsdata::sa22011 %>% 
  filter(state_name_2011 == "New South Wales")

## Create centroids set
centroids <- create_centroids(tas_sa2, "sa2_name_2011")

## Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.12,
  ## same size used in create_grid
  hex_filter = 10,
  use_neighbours = tas_sa2,
  focal_points = capital_cities %>%
    filter(points == "Hobart"),
  width = 35,
  verbose = FALSE
)

fort_tas <- tas_sa2 %>%
  fortify_sfc()

## Make a plot
library(ggplot2)
tas_hexmap <- ggplot() +
  geom_polygon(aes(
      x = long,
      y = lat,
      group = interaction(sa2_name_2011, polygon)
    ),
    fill = "white",
    colour = "lightgrey",
    data = fort_tas
  ) +
  geom_polygon(aes(
    x = long,
    y = lat,
    group = hex_id,
    fill = rownumber
  ),
  data = fort_hex) +
  scale_fill_distiller("", palette = "PRGn") +
  coord_equal() + theme_void()
tas_hexmap
```

```{r mapgrid}

###############################################################################

aus_grid <- gridExtra::grid.arrange(aus_ggchoro, aus_ggcont, aus_ggncont, aus_ggdorl, nrow = 2)

aus_grid
ggsave(filename = "figures/aus_grid.png", plot = aus_grid,
  device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)
```

<!-- \begin{figure}[h] -->
<!-- \centering -->
<!-- \includegraphics[width=14cm]{figs/nsw_grid.png} -->
<!-- \caption{\label{fig:nsw_grid}A collection of alternative maps of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Melanoma for males, ranging from much lower than average (blue) to much higher than average (red)} -->
<!-- \end{figure} -->

<!-- contiguous -->
When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS].
The impact of a disease over connected communities can be seen by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in figure \ref{fig:nsw_grid}b) draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS]. This shown in the northern SA2s of New South Wales with an average (yellow) estimated SIR of Melanoma for males.

<!-- non-contiguous -->
The non-contiguous cartogram in figure \ref{fig:nsw_grid}c) also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of white space can be meaningful in non-contiguous cartograms [@ECGC], in this example the disparity between large rural areas and small urban areas means that the city of Hobart, shown in the inset map looks reasonable, but these units are not visible in the context of the New South Wales state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

<!-- Dorling -->
The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \ref{fig:nsw_grid}d) shows the New South Wales SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in New South Wales that were not apparent in figure \ref{fig:nsw_grid} a), b) or c).

@@

The following code creates the hexagon tile map for all the Statistical Areas at Level 2 in New South Wales. These steps can be executed by the main function, `create_hexmap`, or can be implemented separately for more flexibility.

@@

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/aus_gghexmap.png}
\caption{\label{fig:melanoma-hex}A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR, they range from much lower than average (blue) to much higher than average (red)}
\end{figure}

@@

The polygon set of Statistical Areas at Level 2 (SA2) [@abs2011] of New South Wales in 2011 is provided with the `sugarbag` package as `tas_sa2`. 
A single column of the data set is used to identify the unique areas. 
In this case, the unique SA2 names for each SA2 have been used.

The longitude and latitude centre of the capital cities of Australia are used as focal points to allocate each geographic area around the closest capital city. Hobart will be the common focal point, as this example uses only the areas in the state of New South Wales.

@@

<!-- ```{r grid2, fig.cap = "Grid points to create a tile map.", out.width = "100%", echo = FALSE} -->
<!-- knitr::include_graphics(path = "figures/2grid.png") -->
<!-- ``` -->

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/2grid.png}
\caption{\label{fig:grid2}Grid points to create a hexagon tile map.}
\end{figure}

@@

The grid points are kept only if they fall between the rolling average of the minimum and maximum centroid values after accounting for the buffer distance, for each row and column of the grid. The sparsely populated South-West region of National Park has much fewer points available compared to the South-East region containing the city of Hobart.

<!-- ```{r filter-grid, fig.cap = "All possible hexagon locations from the initial grid are shown with blue outlines. The blue dots show the grid points left to choose from after the buffer step. The rolling windows show the collections of rows and columns used to filter the hexagon locations.", echo = FALSE, out.width = "100%"} -->
<!-- knitr::include_graphics(path = "figures/3grid.png", dpi = 300) -->
<!-- ``` -->

\begin{figure}[h]
\centering
\includegraphics[width=10cm]{figs/3grid.png}
\caption{\label{fig:filter-grid}The remaining hexagon locations from all the possible points in the initial grid after the buffer is applied. The blue dots show the grid points left to choose from after the buffer step.}
\end{figure}

@@

<!-- ```{r buffers, fig.cap = "Filter for grid points within a square, then circular, distance for those closest to the centroid.", echo = FALSE, out.width = "100%"} -->
<!-- knitr::include_graphics(path = "figures/4grid.png") -->
<!-- ``` -->

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/4grid.png}
\caption{\label{fig:buffers}The remaining available hexagon locations after filtering for the grid points within a square distance, then circular around the centroid.}
\end{figure}

@@

<!-- ```{r buffs, fig.cap = "A complete hexagon tile map of New South Wales.", out.width = "100%", echo = FALSE} -->
<!-- knitr::include_graphics(path = "figures/6allocate.png") -->
<!-- ``` -->

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/6allocate.png}
\caption{\label{fig:buffs}A complete hexagon tile map of New South Wales.}
\end{figure}

A complete hexagon tile map of New South Wales is created by applying the algorithm steps to each centroid.
The hexagon tile map visualisation is used below to visualise the Australian Cancer Atlas data [@TACA]. Two views of the same data are produced by filling according to the thyroid Cancer Standardised Incidence Rates (SIRs) downloaded from the Australian Cancer Atlas site. This small example in Figure \ref{fig:sir} shows the group of blue areas in the Hobart CBD more prominently in the hexagon tile map (b).
The small red areas visible in the choropleth map (a) along the north coast are much larger in the hexagon tile maps. The hexagon tile map shows less yellow, this no longer overwhelms the map space with the information regarding the rural areas.

@@

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/7SIR.png}
\caption{\label{fig:sir}The Australian Cancer Atlas data displayed on a choropleth and hexagon tile map. The colour of each Statistical Area of Australian at Level 2 coloured by SIR. A choropleth map (a) of SIR is paired with a hexagon tile map (b) to contrast the colours that are made obvious when every SA2 is equally represented.}
\end{figure}

@@

The `gganimate` [@gganimate] package can be used to make an animation.
It requires connecting the polygons for each area in two displays, which can be done using the `sf_id` variable, such as the SA2 name.
The animation connecting these two displays will highlight the rapid growth of the inner-city areas that emphasises the density of the communities. The hexagons that move the furthest will move rapidly in the animation. The rapid decreases of the large rural areas also show how much greater the landmass of Statistical Areas can be.

@@

Alternative maps highlight the value of the statistics across the geographic units. Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing each administrative area. This acknowledges that the amount of residents can be different but recognises that each population area is equally important. The solution to this visualisation problem has equally sized areas, with neighbourhood boundary connections. This map algorithm is implemented in the `sugarbag` [@sugarbag] package written for `R`.
The `sugarbag` package creates tessellated hexagon tile maps. The Australian application preserves the spatial relationships, emphasising capital cities. The hexagon tile map is a visualisation solution that highlights spatial distributions.

These hexagons equally represent each area. However, the tessellation does not allow the size of the hexagons to represent another variable, similar to the choropleth maps.
The algorithm is heavily dependent on the focal points used, as this determines the order of allocation. It works on the assumption that viewers can use directional relationships to identify their neighbourhoods but this can be aided by the animation.
With careful consideration of the choropleth map, the small geographic inner city areas may have been noticed by viewers, but the hexagon tile map display emphasises them. The communities in northern New South Wales and the Northern territory do not draw attention because of their size as in the choropleth, but their colour is still noticeably below average when contrasted with the hexagons further south.

<!-- why hexagon tile map addresses deficiencies -->
The Dorling cartogram was the most influential in the design of the hexagon tile map.
The use of a hexagon provides a simple shape that shifts the focus of the visualisation from the shape of SA2s to the distribution effects of a disease on communities.  not impeded by the irregular and unusual shapes created by a contiguous cartogram and the immense white space created on the non-contiguous cartogram.
Hexagons can emphasise the relationship between neighbouring communities by allowing their sides to touch. The tessellation crated by adjoining hexagons for densely populated areas is helpful as it shows that most communities are not islands onto themselves, but are intrinsically connected to those around them. As seen in the Dorling cartogram in figure \ref{fig:nsw_grid}c), the tessellation that allows boundary connections was not imposed for rural communities, where centroid of the geographic unit was located far from others, a feature that allows distant but dense populations of the coastal towns of New South Wales to be recognised.
